---
title: "Assignment 1"
author: "Jennifer Horowitz"
date: "9/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warnings = FALSE, messages=FALSE)
```


```{r load libraries}
library(tidyverse)
library(Lahman)
library(kableExtra)
```


```{r load data}
gdp<- read.csv('API_NY.GDP.PCAP.CD_DS2_en_csv_v2_1429392 (1).csv',skip = 4) %>% 
    select(Country.Name, X1996) %>% 
    rename(
        gdp_per_capita = X1996
    )
```

# Data

I decided to explore a subset of baseball data for batters in 1996. My unit level for analysis is at the player level. Data is from the [Lahman database](http://www.seanlahman.com/baseball-archive/statistics/) installed via the Lahman package.

The second set of data used is Gross Domestic Product (GDP) per capita from 1996, downloaded from the [World Bank](https://data.worldbank.org/indicator/NY.GDP.PCAP.CD?view=chart). Data is adjusted to current United States dollars (USD).

# Variables Chosen

I chose variables:

1. nameFirst: First name of the player

2. nameLast: Last name of the player

3. birthDate: Date player was born

4. birthCountry: Country the player was born

5. weight: Weight of the player (in pounds)

6. height: Height of the player (in inches)

7. bats: If the player bats right or left handed

8. batting_average: Batting average of the player. Calculated by their number of hits divided by at bats

9. slugging_average: Average slugging percentage. Total number of bases dived by at bats.

10. gdp_per_capita: GDP from 1996, normalized to 2019 USD

```{r create dataset}
batting_stats_1996 <- Batting %>% 
    filter(yearID == 1996 & AB > 0) %>% 
    inner_join(People) %>% 
    mutate(
        batting_average = H/AB,
        singles = H - (X2B +X3B + HR),
        total_bases = (1 * singles + 2 * X2B + 3 * X3B + 4 * HR) / AB,
        slugging_average = total_bases/AB
    ) %>% 
    select(nameFirst, nameLast, birthDate,birthCountry,
           weight, height, bats, batting_average, slugging_average
           ) %>% 
    mutate(birthCountry = str_replace_all(birthCountry, 'USA', 'United States')) %>% 
    inner_join(gdp, by = c('birthCountry' = 'Country.Name'))
```

Below is an example of the dataset:

```{r}
batting_stats_1996 %>% 
    head(10) %>% 
    kable() %>% 
    kableExtra::kable_styling()
```

My final dataset ends up with 10 variables and 757 players.

# Save the data for future use

```{r}
 #write.csv(batting_stats_1996,'../Desktop/GitHub/jhorowitz-quant/baseballstats.csv')
```

# Types of variables


### Continuous Variables

Batting average, slugging average, height, weight, GDP per capita

### Categorical Variables 

First name, last name, birth country, bath dates 

### Ordinal Variables 

Birth date 

# Research question

How can batting average be explained with the selected variables? 




