---
title: "Assignment 4"
author: "Jennifer Horowitz"
date: "9/30/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning = F, message = F)

options(scipen = 999)
```


```{r}
library(tidyverse)
library(jtools)
library(interactions)
library(ggthemes)
```

```{r}
baseball <- read.csv('baseballtidy.csv') 

baseball <- baseball %>% 
select(-X)
```


### Linear Regression


#### Model 1


The model shows that there is a positive relationship between weight and batting average. For every pound increase in weight, the batting average increase 0.0003. Weight has a p value of 0.21 which is not statistically significant. 

Height is statistically significant. Every inch in height decreases batting average by .008130894. 

Bats left is not statistically significant. If you bat left your batting average would go down .01 points. 

Bats right is statistically significant. If you bat right your batting average would down 0.04 points. 

Slugging average is statistically significant. If your slugging average is positive then your batting average goes up .9. 

GDP per capita is not statistically significant. If your GDP per capita goes up then batting average goes down by a very small margin. 

North American-born players are not statistically significant. If you're from North America your batting average goes up .05. 

Players born in Europe and Asia were not statistically significant. If you are from Europe/Asia then your batting average goes up .06. 

Players from South America were not statistically significant. If you are from South America then your batting average goes up .12. 


My adjusted R-squared . 3482. It only explains 35% of the variance. 

```{r}
model1 <- lm(batting_average~.,data=baseball)
summary(model1)
```
### Model 2 

#### Interaction Term with GDP Per Term Continent 


GDP per capita is often tied to what continent it relates to so the interaction term I will use is GDP per capita and continent

My adjusted R squared was .3466. Which was slightly worse than model 1. 


```{r}
model2 <- lm(batting_average~.+ gdp_per_capita:continent,data=baseball)
summary(model2)
```
### Model 3

#### Using only the 3 significant variables. 

Only the models that were significant in model 1. 

My adjusted R squared is .3493 which is my best model so far. 

```{r}
model3 <- lm(batting_average~ height + bats + slugging_average,data=baseball)
summary(model3)
```



### Visualizing the results 

```{r}
ModelFit <- tibble(model = c(1, 2, 3),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared,
                                summary(model3)$adj.r.squared))
ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 3, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_solarized()
```

Model 3 is the best of the models based on the R-squared. 