---
title: "Assignment 3"
author: "Jennifer Horowitz"
date: "9/22/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
```{r, message=FALSE, echo=FALSE}
library(tidyverse)
library(ggplot2)
library(MASS)
library(ggthemes)
library(wesanderson)
```
```{r}
baseball<-read_csv('baseballstats.csv')
```

```{r}
baseball<-baseball %>% 
  mutate(
    continent = case_when (
      birthCountry =="United States"~"North America",
      birthCountry =="Jamaica"~"North America",
      birthCountry =="Nicaragua"~"Central America",
      birthCountry =="Mexico"~"Central America",
      birthCountry =="Panama"~"Central America",
      birthCountry =="Cuba"~"Central America",
      birthCountry =="Curacao"~"South America",
      birthCountry =="Colombia"~"South America",
      TRUE ~ "other"
    )
  )
```
**GDP per capita-- removed 

### Continuous Variables 

#### Weight and Height



I start by comparing weight and height. 

``` {r}
corweightheight <- with(baseball,cor.test(weight,height))
corweightheight
```

We can visualize the same relationship with a scatter plot. Weight and height are positively correlated. Pretty strong positive correlation between weight and height which is also statistically significant. Typically the taller people get the heavier they are. 

I'll also add a trendline to the scatterplot using geom_smooth().

``` {r}
ggplot(baseball, aes(x = weight, y = height)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  scale_x_continuous(name = "Baseball Player Weight") +
  scale_y_continuous(name = "Baseball Player Height") +
                     #breaks = seq(0, 6000, by = 500),
                    # labels = paste("$", 
                                    #prettyNum(seq(0, 6000, by = 500),
                                              #big.mark = ","))) +
  annotate(geom = "text", x = 0, y = 84, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corweightheight$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corweightheight$conf.int[2], digits = 2)),
           hjust = 0) 
  
```



#### Weight and Batting Average 



``` {r}
corweightandbatting <- with(baseball,cor.test(weight,batting_average))
corweightandbatting
```

``` {r}
ggplot(baseball, aes(x = weight, y = batting_average)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_clean() +
  scale_x_continuous(name = "Baseball Player Weight") +
  scale_y_continuous(name = "Baseball Player Batting Average") +
  annotate(geom = "text", x = 0, y = .8, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corweightandbatting$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corweightandbatting$conf.int[2], digits = 2)),
           hjust = 0)
  
```    

Weight and batting average have almost no linear relationship. At the 95% confidence interval it exhibits no relationship-- it's not statistically signifcant. The p value = .5. 

#### Weight and Slugging Average

The two variables experience no linear relationship but it is statistically signifiant. 

``` {r}
corweightandslugging <- with(baseball,cor.test(weight,slugging_average))
corweightandslugging
```

``` {r}
ggplot(baseball, aes(x = weight, y = slugging_average)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_dark() +
  scale_x_continuous(name = "Baseball Player Weight") +
  scale_y_continuous(name = "Baseball Player Slugging Average") +
  annotate(geom = "text", x = 0, y = .5, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corweightandslugging$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corweightandslugging$conf.int[2], digits = 2)),
           hjust = 0)
  
```  

#### Height and Slugging Average


The two variables exhibit no linear relationship and it is statistically significant. 

``` {r}
corheightandslugging <- with(baseball,cor.test(height,slugging_average))
corheightandslugging
```

``` {r}
baseball %>% 
 filter(slugging_average<=0.4) %>% 
  ggplot(aes(x = height, y = slugging_average)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  xlim(65,84) +
  scale_x_continuous(name = "Baseball Player Height") +
  scale_y_continuous(name = "Baseball Player Slugging Average") +
  annotate(geom = "text", x = 69, y = .2, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corheightandslugging$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corheightandslugging$conf.int[2], digits = 2)),
           hjust = 0)
 
```


#### Height and Batting Average 

There is no linear relationship between these two variables. It is barely not statistically significant. 



``` {r}
corheightandbatting <- with(baseball,cor.test(height,batting_average))
corheightandbatting
```

``` {r}
ggplot(baseball, aes(x = height, y = slugging_average)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_bw() +
  xlim(65,84) +
  scale_x_continuous(name = "Baseball Player Height") +
  scale_y_continuous(name = "Baseball Player Slugging Average") +
  annotate(geom = "text", x = 70, y = .1, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corheightandslugging$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corheightandslugging$conf.int[2], digits = 2)),
           hjust = 0)
  
```

#### Slugging Average and Batting Average 

``` {r}
corsaandba<- with(baseball,cor.test(slugging_average,batting_average))
corsaandba
```
``` {r}
ggplot(baseball, aes(x = batting_average, y = slugging_average)) +
  geom_point(size = 0.5, alpha = 0.25, position = "jitter") +
  geom_smooth() +
  theme_economist_white() +
  scale_x_continuous(name = "Batting Average") +
  scale_y_continuous(name = "Slugging Average") +
  xlim(0,0.5) + 
  ylim(0,0.1) +
  annotate(geom = "text", x = 0, y = .1, 
           label = paste("Correlation (95-percent confidence):",
                         prettyNum(corsaandba$conf.int[1], digits = 2), 
                         "to",
                         prettyNum(corsaandba$conf.int[2], digits = 2)),
           hjust = 0)
  
```

### Continuous Variable vs. Categorical Variable (3+ levels)
 
#### Weight vs. Bats 

There is a significant association between weight and bats. 

``` {r}
weightbatsanova <- aov(weight ~ bats, data = baseball)
summary(weightbatsanova)
```


``` {r}
weightbatsdif <- TukeyHSD(weightbatsanova)
as_tibble(cbind(pair = row.names(weightbatsdif$bats),
               weightbatsdif$bats))
```

The graph demonstrates that there is a higher preponderance of right handed players and the biggest difference lies between those that are left-handed and those who are "switch hitters".

``` {r}
ggplot(baseball, aes(x =bats, y = weight)) +
  geom_boxplot() +
  theme_bw() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Which Side of the Plate You Bat From") +
  scale_y_continuous(name = "Weight") 

```


#### Weight vs. Continents  

There is not a significant linear relationship between weight and continent. 

``` {r}
weightcontinentanova <- aov(weight ~ continent, data = baseball)
summary(weightcontinentanova)
```



``` {r}
weightcontinentdif <- TukeyHSD(weightcontinentanova)
as_tibble(cbind(pair = row.names(weightcontinentdif$continent),
               weightcontinentdif$continent))
```

The heaviest players originated from Central and South America. 

``` {r}
ggplot(baseball, aes(x = continent, y = weight)) +
  geom_boxplot() +
  theme_economist() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Which Continent Player Comes From") +
  scale_y_continuous(name = "Weight") 

```

#### Height and Bats 

There is a significant association between height and bats. 

``` {r}
heightbatsanova <- aov(height ~ bats, data = baseball)
summary(heightbatsanova)
```


``` {r}
heightbatsdif <- TukeyHSD(heightbatsanova)
as_tibble(cbind(pair = row.names(heightbatsdif$bats),
               heightbatsdif$bats))
```
The people that bat from both sides are on the shorter end of the height data available.


``` {r}
ggplot(baseball, aes(x =bats, y = height)) +
  geom_boxplot() +
  theme_fivethirtyeight() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Which Side of the Plate You Bat From") +
  scale_y_continuous(name = "Height") 

```


#### Height and Continent

There is not a significant relationship between the two variables. 

``` {r}
heightcontinentanova <- aov(height ~ continent, data = baseball)
summary(heightcontinentanova)
```



``` {r}
heightcontinentdif <- TukeyHSD(heightcontinentanova)
as_tibble(cbind(pair = row.names(heightcontinentdif$continent),
               heightcontinentdif$continent))
```


There was no significant relationship between the two variables. 


``` {r}
ggplot(baseball, aes(x = continent, y = height)) +
  geom_boxplot() +
  theme_solarized() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Which Continent Player Comes From") +
  scale_y_continuous(name = "Height") 

```

South America-other experienced the most significant difference. 

#### Batting Average and Bats


There is a significant association between batting average and which side of the plate players hit from. 

``` {r}
babatsanova <- aov(batting_average ~ bats, data = baseball)
summary(babatsanova)
```


``` {r}
babatsdif <- TukeyHSD(babatsanova)
as_tibble(cbind(pair = row.names(babatsdif$bats),
               babatsdif$bats))
```

The magnitude on the for players to hit from the right-hand side is higher. The biggest range is between the players who hit from the left and players who hit from both sides. 


``` {r}
ggplot(baseball, aes(x = bats, y = batting_average)) +
  geom_boxplot() +
  theme_wsj() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Side of the Plate Players Bat On") +
  scale_y_continuous(name = "Batting Average") 

```



#### Batting Average and Continent 

There is not a significant linear relationship between batting average and continent. 

``` {r}
bacontinentanova <- aov(batting_average ~ continent, data = baseball)
summary(bacontinentanova)
```



South America-other has the biggest difference (range) between them. 



``` {r}
bacontinentdif <- TukeyHSD(bacontinentanova)
as_tibble(cbind(pair = row.names(bacontinentdif$continent),
             bacontinentdif$continent))
```

``` {r}
ggplot(baseball, aes(x = continent, y = batting_average)) +
  geom_boxplot() +
  theme_stata() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Continent of Origin") +
  scale_y_continuous(name = "Batting Average") 
```

#### Continent and Slugging Average 

There is not a significant linear relationship between slugging average and continent. 

``` {r}
sacontinentanova <- aov(slugging_average ~ continent, data = baseball)
summary(sacontinentanova)
```



South America-other has the biggest difference (range) between them. 


``` {r}
sacontinentdif <- TukeyHSD(sacontinentanova)
as_tibble(cbind(pair = row.names(sacontinentdif$continent),
             sacontinentdif$continent))
```
``` {r}
ggplot(baseball, aes(x = continent, y = slugging_average)) +
  geom_boxplot() +
  theme_classic() +
  theme(axis.text.x = element_text(angle = 90)) +
  scale_x_discrete(name = "Continent of Origin") +
  scale_y_continuous(name = "Slugging Average") 
```



### Chi Square Test: Two Categorical Variables 

#### Bats and Continents 

The linear relationship between bats and continents is not significant. 

``` {r chi square}
chi_sq <- chisq.test(baseball$bats, baseball$continent)
  
chi_sq
```
``` {r chi square observed}
chi_sq$observed
```

Here are the values we would expect if the two variables were not related:

``` {r chi square expected}
chi_sq$expected
```