---
title: "Assignment 5"
author: "Jennifer Horowitz"
date: "10/9/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
```{r}

```



```{r warning=FALSE, message=FALSE}
library(tidyverse)
library(jtools)
library(interactions)
library(ggthemes)
```

```{r}
baseball <- read.csv('baseballtidy.csv') 

baseball <- baseball %>% 
select(-X)
```


### Linear Regression


#### Model 1

Log transform of slugging average because it isn't approximately normal and will use the interaction term weight and height becauase weight usually increases with height. 

To do the log transformation you need to remove all 0s from the slugging average.



```{r}
baseball <- baseball %>% 
  filter(slugging_average>0)



model1 <- lm(batting_average~weight + height + bats + log(slugging_average) + weight:height,data=baseball) 
summary(model1)
```


This model has been the worst of of the models I have run so far. The adjusted R square value was .08 which was far worse than than any of the other adjusted R square values of previous models I have run.  


### Model 2 

I next logged GDP per capita to normalize it. The adjusted R square value was .09. It, similar to model 1, preformed badly. 

```{r}
model2 <- lm(batting_average~weight + height + bats + log(gdp_per_capita) + continent, data=baseball) 
summary(model2) 
```




### Visualizing the results 

The best model I ran of the models here was model 1. 

```{r}
ModelFit <- tibble(model = c(1, 2),
                   R_square = c(summary(model1)$adj.r.squared,
                                summary(model2)$adj.r.squared))
ggplot(ModelFit, aes(x = model, y = R_square)) +
  geom_line() +
  scale_x_continuous(name = "",
                   breaks = breaks <- seq(1, 2, by = 1),
                   labels = paste("Model", breaks)) +
  scale_y_continuous(name = "Adjusted R-squared value") +
  theme_solarized()
```